# R package 

## Introduction {-}

I created a R package to support my portfolio, following [The Whole Game demo](https://r-pkgs.org/whole-game.html) made by Hadley Wickham. The demo guided me through setting up my own package and helped me create something useful on my own.

----

## Copopa {-}

To start a new package, a name is necessary. I combined my name, "portfolio" and "package" together into a fun word and checked availability with the {available} package. 

The name was available, so I set up the project and the github repository to get started.

A package to support my portfolio would reduce duplicated code. But scanning through my portfolio, I noticed that there wasn't a lot of code or duplicated code. Luckily, I found some that would work for this. The package that I made contains four functions, which I will elaborate following the next example dataset:

```{r example data}

# Create data frame
name <- c("Noa", "Luke", "Oliver", "Anna", "Julie")
gender <- c("F", "M", "M", "F", "F")
age <- c(21, 29, 27, 29, 24)
test <- data.frame(name, gender, age)

kbl(test, caption = "Table 2: Example data frame for copopa package") %>% 
  kable_paper(bootstrap_options = "striped", full_width = F)

```

### multi_check() {-}

This functions checks the data types of up to three datasets, or columns in it. Full sets can be used as input, or in this case columns.

```{r multi check}

multi_check(test$name, test$gender, test$age)

```

### mima() {-}

This function calculates the minimum and maximum of data. The data must be numeric.

```{r mima}

mima(test$age)

```

### mean_sd() {-}

This function calculates the mean and standard deviation of data. The data must be numeric.

```{r mean sd}

mean_sd(test$age)

```

### save_to_csv_rds() {-}

This function saves a data frame to a CSV file and a RSD file. Fill in target data set and the name of the wanted output.

```{r save, eval = FALSE}

save_to_csv_rds(test, "test")

```

For the manual and contents of the package, please visit [the copopa github repository](https://github.com/coconiemel/copopa).

## Install and load {-}

```{r install and load package, eval = FALSE}

# Only if not obtained already
install.packages("devtools")
library(devtools)

# Installation
devtools::install_github("coconiemel/copopa")

# Load
library(copopa)

```



